The function--sheaf dictionary for smooth commutative group schemes over finite fields

### 2014-08-10 dictionary.tex
Moving sections around and removing references to flabby quasicharacter sheaves

%%%%%%%%%%%% MISCELLANEOUS OPERATORS %%%%%%%%%%%%
\newcommand{\sheafHom}{{\mathscr{H}\hskip-4pt{\it o}\hskip-2pt{\it m}}}
\newcommand{\abs}[1]{{\vert #1 \vert}}
\newcommand{\ceq}{{\, :=\, }}
\newcommand{\tq}{{\ \vert\ }}
\newcommand{\iso}{{\ \cong\ }}
%% Limits
\newcommand{\invlim}[1]{\lim\limits_{\overleftarrow{#1}}}
\newcommand{\dirlim}[1]{\lim\limits_{\overrightarrow{#1}}}
\newcommand{\limit}[1]{\mathop{\textsc{lim}}\limits_{#1}}
\newcommand{\colimit}[1]{\mathop{\textsc{colim}}\limits_{#1}}
%% Fonts for quasicharacter sheaves
\newcommand{\qcs}[1]{{\mathcal{#1}}}
\newcommand{\gqcs}[1]{{\mathcal{\bar #1}}}
\newcommand{\dualgcs}[1]{\gqcs{#1}^\dagger}
\newcommand{\dualcs}[1]{\qcs{#1}^\dagger}
%% Categories
\newcommand{\QC}{{\mathcal{Q\hskip-0.8pt C}}}
\newcommand{\lQC}{{\QC_\text{flb}}}
\newcommand{\lQCiso}[1]{\lQC(#1)_{/\text{iso}}}
\newcommand{\rQC}{{\QC_{\text{rig}}}}
\newcommand{\rbQC}{{\QC_0}}
\newcommand{\rfQC}{{\QC_\text{fin}}}
\newcommand{\rQCiso}[1]{\rQC(#1)_{/\text{iso}}}
\newcommand{\rbQCiso}[1]{\rbQC(#1)_{/\text{iso}}}
\newcommand{\rfQCiso}[1]{\rfQC(#1)_{/\text{iso}}}
\newcommand{\trFrob}[1]{t_{#1}}
\newcommand{\TrFrob}[1]{T_{#1}}
%% amsmath replacements for \atop
\newcommand{\latop}[2]{\genfrac{}{}{0pt}{0}{#1}{#2}}
\newcommand{\partop}[2]{\genfrac{(}{)}{0pt}{0}{#1}{#2}}
%% Labeled items
\makeatletter
\newcommand{\labitem}[2]{
\def\@itemlabel{\textbf{#1}}
\item
\def\@currentlabel{#1}\label{#2}}
\makeatother
%% Shorthand for bars
\renewcommand{\bf}{\bar{f}}
\newcommand{\bg}{\bar{g}}
\newcommand{\bm}{\bar{m}}
\newcommand{\bG}{\bar{G}}
\newcommand{\bH}{\bar{H}}
%% Spacing control
\newcommand{\tight}[3]{\hspace{-#1pt}{#2}\hspace{-#3pt}}
\newcommand{\GxG}{\text{$G \tight{1}{\times}{1} G$}}
\newcommand{\bGxG}{\text{$\bar{G} \tight{1}{\times}{1} \bar{G}$}}
\newcommand{\bfxf}{\text{$\bar{f} \tight{1}{\times}{1} \bar{f}$}}
\newcommand{\GxxG}{\text{$G \tight{1}{\times}{1} G$}}
\newcommand{\LxL}{\text{$\gqcs{L} \tight{0}{\boxtimes}{0} \gqcs{L}$}}
\newcommand{\ExE}{\text{$\qcs{E}\tight{0}{\boxtimes}{0}\qcs{E}$}}
\newcommand{\bExE}{\text{$\gqcs{E}\tight{0}{\boxtimes}{0}\gqcs{E}$}}
\newcommand{\AxA}{\text{$A \tight{1}{\times}{1} A$}}
\newcommand{\BxB}{\text{$B \tight{1}{\times}{1} B$}}
\newcommand{\GzxGz}{\text{$G^0 \tight{1}{\times}{1} G^0$}}
%% Margin notes
\newcommand\Clifton[1]{\marginpar{\smaller\smaller CC: #1}}
\newcommand\David[1]{\marginpar{\smaller\smaller DR: #1}}
%% Hyphenation override
\hyphenation{quasi-character}

@@

%In order to describe the main result of this paper with more precision,
Let us briefly review how Deligne's function--sheaf dictionary for connected commutative algebraic groups over finite fields \cite{deligne:SGA4.5}*{sommes trig.} translates characters into local systems.
Let $\Fq$ be a finite field and let $G$ be a connected commutative algebraic group over $\Fq$; fix an algebraic closure $\bFq$ of $\Fq$
and let $\bG$ be the base change of $G$ to $\bFq$.
Fix a prime $\ell$ invertible in $\Fq$ and let $\chi : G(\Fq) \to \EEx$ be any character.
Using little more than the Lang map for $G$, $\chi$ determines an $\ell$-adic local system $\qcs{L}$ on $G$ from which the character $\chi$ can be recovered using the Frobenius automorphism of $\bG$.
While this recovery process applies to all $\ell$-adic local systems $\qcs{L}$ on $G$, those that produce characters of $G(\Fq)$ are distinguished by a simple property: there \emph{exists} an isomorphism
\begin{equation}\label{introbox}
m^* \qcs{L} \iso \qcs{L} \boxtimes \qcs{L}
\end{equation}
 of $\ell$-adic local systems on $G$, where $m : G\times G \to G$ is the multiplication map for $G$. The function--sheaf dictionary for $G$ identifies a monoidal category of $\ell$-adic local systems on $G$ such that the group of isomorphism classes in this category is canonically identified with the $\ell$-adic character group of $G(\Fq)$, thus providing the geometrization and categorification of characters of $G(\Fq)$.
%
Although hidden in this discussion, the connectedness of $G$ played a crucial role in the geometrization and categorification of characters of $G(\Fq)$.
In particular, when the isomorphism \eqref{introbox} exists for $\qcs{L}$, it is essentially unique.

Now let $G$ be any smooth commutative group scheme over $\Fq$.
Two new features must be addressed in order to geometrize characters of $G(\Fq)$.
First, we must pass from $\ell$-adic local systems on $G$ to Weil local systems on $G$
and, second, we must keep track of the isomorphisms \eqref{introbox}.
Both issues are addressed in the main definition of this paper:
A \emph{rigid quasicharacter sheaf} on $G$ is a triple $\qcs{L}\ceq
(\gqcs{L},\mu,\phi)$ where $\gqcs{L}$ is an $\ell$-adic local system on $\bG$ of rank~one and $\mu: m^*
\gqcs{L} \to \LxL$ and $\phi : \Frob{G}^*\gqcs{L} \to \gqcs{L}$ are isomorphisms satisfying natural compatibility conditions; here $\Frob{G}$ is the Frobenius automorphism of ${\bar G}$.
A \emph{flabby quasicharacter sheaf} on $G$ is a Weil sheaf $(\gqcs{L},\phi)$ on $G$
for which $(\gqcs{L},\mu,\phi)$ is a rigid quasicharacter sheaf on $G$, for some $\mu$.
The difference between these two notions is detected by $\Hh^2(\pi_0({\bar G}),\EEx)$, where $\pi_0({\bar G})$ is the component group for ${\bar G}$.
%namely, when $\Hh^2(\pi_0({\bar G}),\EEx)$ is non-trivial, if $(\gqcs{L},\phi)$ is a quasicharacter sheaf on $G$ then there are many $\mu$ for which $(\gqcs{L},\mu,\phi)$ is a rigid quasicharacter sheaf on $G$;
%
 In the main result of this paper, Theorem~\ref{thm:snake},
we form a rigid monoidal category $\lQC(G)$ of flabby quasicharacter sheaves on $G$
and show that the Trace of Frobenius provides a functorial group isomorphism
\begin{equation}
\lQCiso{G} \to \Hom(G(\Fq), \EEx).
\end{equation}

@@

%\subsection*{Structure of the paper}

We close the introduction with a brief description of the structure of this paper.
After setting some basic notation in Section~\ref{sec:notation},
in Section~\ref{sec:category} we define rigid quasicharacter sheaves and give an interpretation of
them as Weil sheaves equipped with an isomorphism of the form \eqref{introbox}.
Section~\ref{sec:Frob} introduces Trace of Frobenius,
one direction of the function--sheaf dictionary.
We define pullbacks and products of rigid quasicharacter sheaves in Section~\ref{sec:pullback}, and use
them to prove a functoriality result on Trace of Frobenius.
In Section~\ref{sec:basechange} we study how rigid quasicharacter sheaves behave under base change
and Weil restriction.
Sections~\ref{sec:bounded} and \ref{sec:finite} describe bounded quasicharacter sheaves and finite quasicharacter sheaves, respectively,
forming two important subcategories of $\rQC(G)$.
%These subcategories play a key role in the proof of Proposition~\ref{prop:restriction}.
In Section \ref{sec:connected} we show that all rigid quasicharacter sheaves on $G$ are finite quasicharacter sheaves when $G$ is connected, and that Trace of Frobenius is an isomorphism in this case.
We shift to \'etale group schemes $G$ in Section~\ref{sec:etale} and find a cohomological interpretation
of rigid quasicharacter sheaves in this case.
In Section~\ref{sec:restriction} we show that for non-connected $G$, restriction to the identity component
induces an essentially surjective functor of rigid quasicharacter sheaves.
Section~\ref{sec:snake} presents the main theorem of this paper: Trace of Frobenius
is a functorial isomorphism from isomorphism classes of flabby quasicharacter sheaves on $G$
to characters of $G(\Fq)$.

Comments about Joseph Bernstein's comments?

@@

Throughout this paper, $G$ is a smooth, commutative group scheme
over a finite field $\Fq$, and $m : \GxG\to G$ is its multiplication morphism.
%For the sake of brevity, we define a \emph{$\rQC$-scheme} to be such a $G$ with the additional property that the geometric component group of $G$ is finitely generated.  This last condition is only needed in the proof of Lemma \ref{lemma:section}.  A morphism of $\rQC$-schemes is a morphism of group schemes over $\Fq$.

@@

\subsection{Rigid quasicharacter sheaves}\label{sec:category}

\begin{definition}\label{def:QC}
A \emph{rigid quasicharacter sheaf on $G$} is a triple
$\qcs{L}\ceq (\gqcs{L},\mu,\phi)$ where:
\begin{enumerate}
\labitem{(QC.1)}{QC.1} $\gqcs{L}$ is a rank~one $\ell$-adic local system on $\bG$, by which we mean
a constructible $\ell$-adic sheaf on $\bG$, locally constant on each connected
component of $\bG$, whose stalks are one-dimensional $\EE$-vector spaces;
\labitem{(QC.2)}{QC.2} $\mu: \bm^* \gqcs{L} \to \LxL$ is an isomorphism of
sheaves on $\bGxG$ such that the following diagram commutes,
  where $m_3 \ceq m\circ (m\tight{1}{\times}{2}\id) = m\circ (\id\tight{2}{\times}{1} m)$;
  \[
  \begin{tikzcd}[row sep=30]
  \bm_3^*\gqcs{L} \arrow{rr}{(\bm \tight{1}{\times}{2} \id)^*\mu} \arrow[swap]{d}{(\id \tight{2}{\times}{1} \bm)^*\mu}
    &&  \bm^*\gqcs{L} \boxtimes \gqcs{L} \dar{\mu \tight{0}{\boxtimes}{1} \id} \\
    \gqcs{L} \boxtimes \bm^* \gqcs{L} \arrow{rr}{\id \boxtimes \mu}
    &&  \gqcs{L} \tight{0}{\boxtimes}{0} \LxL
  \end{tikzcd}
  \]
\labitem{(QC.3)}{QC.3} $\phi : \Frob{G}^* \gqcs{L} \to \gqcs{L}$ is an
  isomorphism of constructible $\ell$-adic sheaves on $\bG$ compatible with
  $\mu$ in the sense that the following diagram commutes.
  \[
  \begin{tikzcd}[row sep=20]
  \Frob{\GxxG}^* \bm^* \gqcs{L} \arrow{rr}{\Frob{\GxxG}^*\mu}
    && \Frob{\GxxG}^*(\LxL)\\
    \arrow[equal]{u} \bm^*  \Frob{G}^* \gqcs{L} \arrow[swap]{d}{\bm^* \phi}
    && \Frob{G}^*\gqcs{L}\boxtimes \Frob{G}^*\gqcs{L} \dar{\phi\boxtimes \phi} \arrow[equal]{u} \\
    \bm^*\gqcs{L} \arrow{rr}{\mu}
    && \LxL
  \end{tikzcd}
  \]
\end{enumerate}
\end{definition}

Morphisms in the category of rigid quasicharacter sheaves on $G$, denoted by $\rQC(G)$, are defined in the natural way:
\begin{enumerate}
\labitem{(QC.4)}{QC.4} if $\qcs{L} = (\gqcs{L},\mu,\phi)$ and
  $\qcs{L'} = (\gqcs{L'},\mu',\phi')$ are rigid quasicharacter sheaves on $G$ then
  a morphism $\qcs{L} \to \qcs{L}'$ is a map $\alpha : \gqcs{L} \to \gqcs{L'}$
  of constructible $\ell$-adic sheaves on $\bG$ such that the following diagrams both commute.
  \[
  \begin{tikzcd}[column sep=40]
  \Frob{G}^* \gqcs{L} \rar{\Frob{G}^* \alpha} \arrow[swap]{d}{\phi} & \Frob{G}^* \gqcs{L'} \dar{\phi'}
  & & \arrow[swap]{d}{\mu} m^* \gqcs{L} \rar{m^* \alpha} & m^* \gqcs{L'} \dar{\mu'} \\
  \gqcs{L} \rar{\alpha} & \gqcs{L'}
  & {} & \LxL \rar{\tight{1}{\alpha\boxtimes \alpha}{1}} & \gqcs{L'} \tight{0}{\boxtimes}{0} \gqcs{L'}
  \end{tikzcd}
  \]
\end{enumerate}

The category $\rQC(G)$ of rigid quasicharacter sheaves on $G$ is a rigid monoidal category
\cite{etingof:09a}*{\S1.10} under the tensor product
$\qcs{L} \otimes \qcs{L'} \ceq (\gqcs{L}\otimes\gqcs{L'}, \mu\otimes\mu', \phi\otimes \phi')$,
with duals given by applying the sheaf hom functor
$\sheafHom(\ - \ ,\EE)$.
The category of rigid quasicharacter sheaves on $G$ is not abelian, so $\rQC(G)$
is not a tensor category in the sense of \cite{deligne:02a}*{0.1}.
This rigid monoidal category structure for $\rQC(G)$ gives the set $\rQCiso{G}$
of isomorphism classes in $\rQC(G)$ the structure of a group.

\subsection{Flabby quasicharacter sheaves}\label{sec:loose}

Observe that the rule $(\gqcs{L},\mu,\phi) \mapsto (\gqcs{L},\phi)$ defines a forgetful functor from rigid quasicharacter
sheaves on $G$ to ($\ell$-adic) Weil sheaves on $G$ \cite{deligne:80a}*{Def.~1.1.10 (i)}.
%

@@

\begin{definition}\label{def:lQC}
Let $G$ be a smooth commutative group scheme over $\Fq$.
A \emph{flabby quasicharacter sheaf} on $G$ is a Weil sheaf $(\gqcs{L},\phi)$ on $G$
for which $(\gqcs{L},\mu,\phi)$ is a rigid quasicharacter sheaf on $G$
for some $\mu: \bm^* \gqcs{L} \to \gqcs{L} \boxtimes \gqcs{L}$.
%The resulting objects lie in the essential image of the forgetful functor $(\gqcs{L},\mu,\phi) \mapsto (\gqcs{L},\phi)$ from $\rQC(G)$ to Weil sheaves on $G$.
We form the category of flabby quasicharacter sheaves by declaring that a morphism $(\gqcs{L},\phi) \to (\gqcs{L'},\phi')$ is a sheaf homomorphism $\alpha : \gqcs{L} \to \gqcs{L'}$ for which the two diagrams in Condition~\ref{QC.4} commute, for some choice of $\mu$, $\mu'$ such that $(\gqcs{L},\mu,\phi)$ and $(\gqcs{L'},\mu',\phi')$ are rigid quasicharacter sheaves on $G$.
The category of flabby quasicharacter sheaves on $G$ is denoted by $\lQC(G)$.
\end{definition}

Since objects in $\lQC(G)$ are Weil sheaves and since in the first diagram in Condition~\ref{QC.4} ensures that a morphism in $\lQC(G)$ is a morphism of Weil sheaves, it follows that $\lQC(G)$ is a subcategory of Weil sheaves on $G$.
%
Like $\rQC(G)$, the category $\lQC(G)$ of quasicharacter sheaves on $G$ is rigid monoidal. The functor $\rQC(G) \to \lQC(G)$ is monoidal and essentially surjective but neither full nor faithful, generally.
%

We now introduce a tool which will help us understand isomorphism classes of objects in $\rQCiso{G}$ and $\lQCiso{G}$.

Now suppose $\qcs{L} = (\gqcs{L},\mu,\phi)$ is a rigid quasicharacter sheaf on $G$.
The isomorphism $\mu : \bm^* \gqcs{L} \to \LxL$ together with
\ref{QC.3} guarantee
that the function $\trFrob{(\qcs{L},\phi)} : G(\Fq)\to \EEx$ is a group homomorphism.  
Moreover, this group homomorphism is insensitive to $\phi$ in the sense 
that if $(\gqcs{L},\mu,\phi)$ and $(\gqcs{L},\mu,\phi')$ are rigid quasicharacter sheaves, then $\trFrob{(\qcs{L},\phi)} = \trFrob{(\qcs{L},\phi')}$. \todo{Sketch proof?}
This justifies abbreviating $\trFrob{(\gqcs{L},\phi)}$ to $\trFrob{\qcs{L}}$ in this case.
The diagram below defines two group homomorphisms, both denoted $\TrFrob{G}$ and both called the \emph{Trace of Frobenius} for $G$; we trust that context will remove any ambiguity regarding this notation.
\[
\begin{tikzcd}
\rQCiso{G} \arrow{rr} \arrow{dr}[swap]{\qcs{L} \mapsto t_{\qcs{L}}}{\TrFrob{G}} && \arrow{dl}[swap]{\TrFrob{G}}{(\gqcs{L},\phi) \mapsto t_{\gqcs{L},\phi}} \lQCiso{G} \\
& G(\Fq)^* & 
\end{tikzcd}
\]
As remarked above, the group homomorphism $\rQCiso{G} \to \lQCiso{G}$ is surjective;
we will see that $\rQCiso{G} \to \lQCiso{G}$ need not be injective.
However, in Theorem~\ref{thm:snake}, we will see that the group homomorphism $\rQCiso{G} \to G(\Fq)^*$ is surjective
while the main result of this paper, Theorem~\ref{thm:snake}, 
shows that $\lQCiso{G} \to G(\Fq)^*$ is an isomorphism.

\begin{proposition}\label{prop:functorialG}
Trace of Frobenius $\TrFrob{G} : \rQCiso{G} \to G(\Fq)^*$ is a natural transformation
between the two contravariant additive functors
\begin{align*}
F_1 : G &\mapsto \rQCiso{G} \\
F_2 : G &\mapsto G(\Fq)^*
\end{align*}
from the category of smooth commutative group schemes over $\Fq$ to the category of abelian groups.
Likewise, for $\TrFrob{G} : \lQCiso{G} \to G(\Fq)^*$, {\it mut.\hskip-1pt mut.}
\end{proposition}

The proof of Proposition~\ref{prop:functorialG} will be given at the end of Section~\ref{sec:pullback}.

@@

\begin{proposition}\label{prop:pullback}
  If $f : H\to G$ is a morphism of smooth commutative group schemes over $\Fq$ then
  \begin{align*}
  f^* : \rQC(G) &\to \rQC(H) \\
  (\gqcs{L},\mu,\phi) &\mapsto (\bf^*\gqcs{L},(\bfxf)^*\mu,\bf^*F)
  \end{align*}
  defines a monoidal functor dual to $f \colon H(\Fq) \to G(\Fq)$:
  \[
  \begin{tikzcd}[row sep=20, column sep=30]
   \rQCiso{G} \rar{f^*} \arrow[swap]{d}{\TrFrob{G}} & \rQCiso{H} \dar{\TrFrob{H}} \\
   G(\Fq)^* \rar & H(\Fq)^*
  \end{tikzcd}
  \]
  is a commutative diagram of groups.  Moreover, $(f\circ g)^* = g^* \circ f^*$.
\end{proposition}

@@

We end this section with a simple result on products of rigid quasicharacter sheaves,
which is needed to show that the isomorphism above is functorial.

>>>>

%%%%%%%%%%%% MISCELLANEOUS OPERATORS %%%%%%%%%%%%
\newcommand{\sheafHom}{{\mathscr{H}\hskip-4pt{\it o}\hskip-2pt{\it m}}}
\newcommand{\abs}[1]{{\vert #1 \vert}}
\newcommand{\ceq}{{\, :=\, }}
\newcommand{\tq}{{\ \vert\ }}
\newcommand{\iso}{{\ \cong\ }}
%% Limits
\newcommand{\invlim}[1]{\lim\limits_{\overleftarrow{#1}}}
\newcommand{\dirlim}[1]{\lim\limits_{\overrightarrow{#1}}}
\newcommand{\limit}[1]{\mathop{\textsc{lim}}\limits_{#1}}
\newcommand{\colimit}[1]{\mathop{\textsc{colim}}\limits_{#1}}
%% Fonts for quasicharacter sheaves
\newcommand{\qcs}[1]{{\mathcal{#1}}}
\newcommand{\gqcs}[1]{{\mathcal{\bar #1}}}
\newcommand{\dualgcs}[1]{\gqcs{#1}^\dagger}
\newcommand{\dualcs}[1]{\qcs{#1}^\dagger}
%% Categories
\newcommand{\QC}{{\mathcal{Q\hskip-0.8pt C}}}
\newcommand{\lQC}{{\QC_\text{flb}}}
\newcommand{\lQCiso}[1]{\lQC(#1)_{/\text{iso}}}
\newcommand{\bQC}{{\QC_0}}
\newcommand{\fQC}{{\QC_\text{fin}}}
\newcommand{\QCiso}[1]{\QC(#1)_{/\text{iso}}}
\newcommand{\bQCiso}[1]{\bQC(#1)_{/\text{iso}}}
\newcommand{\fQCiso}[1]{\fQC(#1)_{/\text{iso}}}
\newcommand{\trFrob}[1]{t_{#1}}
\newcommand{\TrFrob}[1]{T_{#1}}
%% amsmath replacements for \atop
\newcommand{\latop}[2]{\genfrac{}{}{0pt}{0}{#1}{#2}}
\newcommand{\partop}[2]{\genfrac{(}{)}{0pt}{0}{#1}{#2}}
%% Labeled items
\makeatletter
\newcommand{\labitem}[2]{
\def\@itemlabel{\textbf{#1}}
\item
\def\@currentlabel{#1}\label{#2}}
\makeatother
%% Shorthand for bars
\renewcommand{\bf}{\bar{f}}
\newcommand{\bg}{\bar{g}}
\newcommand{\bm}{\bar{m}}
\newcommand{\bG}{\bar{G}}
\newcommand{\bH}{\bar{H}}
%% Spacing control
\newcommand{\tight}[3]{\hspace{-#1pt}{#2}\hspace{-#3pt}}
\newcommand{\GxG}{\text{$G \tight{1}{\times}{1} G$}}
\newcommand{\bGxG}{\text{$\bar{G} \tight{1}{\times}{1} \bar{G}$}}
\newcommand{\bfxf}{\text{$\bar{f} \tight{1}{\times}{1} \bar{f}$}}
\newcommand{\GxxG}{\text{$G \tight{1}{\times}{1} G$}}
\newcommand{\LxL}{\text{$\gqcs{L} \tight{0}{\boxtimes}{0} \gqcs{L}$}}
\newcommand{\ExE}{\text{$\qcs{E}\tight{0}{\boxtimes}{0}\qcs{E}$}}
\newcommand{\bExE}{\text{$\gqcs{E}\tight{0}{\boxtimes}{0}\gqcs{E}$}}
\newcommand{\AxA}{\text{$A \tight{1}{\times}{1} A$}}
\newcommand{\BxB}{\text{$B \tight{1}{\times}{1} B$}}
\newcommand{\GzxGz}{\text{$G^0 \tight{1}{\times}{1} G^0$}}
%% Margin notes
\newcommand\Clifton[1]{\marginpar{\smaller\smaller CC: #1}}
\newcommand\David[1]{\marginpar{\smaller\smaller DR: #1}}
%% Hyphenation override
\hyphenation{quasi-character}

@@

Let us briefly review how Deligne's function--sheaf dictionary for connected commutative algebraic groups over finite fields \cite{deligne:SGA4.5}*{sommes trig.} translates characters into local systems.
Let $\Fq$ be a finite field and let $G$ be a connected commutative algebraic group over $\Fq$; fix an algebraic closure $\bFq$ of $\Fq$
and let $\bG$ be the base change of $G$ to $\bFq$.
Fix a prime $\ell$ invertible in $\Fq$ and let $\chi : G(\Fq) \to \EEx$ be any character.
Using little more than the Lang map for $G$, $\chi$ determines an $\ell$-adic local system $\qcs{L}$ on $G$ from which the character $\chi$ can be recovered using the Frobenius automorphism of $\bG$.
While this recovery process applies to all $\ell$-adic local systems $\qcs{L}$ on $G$, those that produce characters of $G(\Fq)$ are distinguished by a simple property: there \emph{exists} an isomorphism
\begin{equation}\label{introbox}
m^* \qcs{L} \iso \qcs{L} \boxtimes \qcs{L}
\end{equation}
 of $\ell$-adic local systems on $G$, where $m : G\times G \to G$ is the multiplication map for $G$. The function--sheaf dictionary for $G$ identifies a monoidal category of $\ell$-adic local systems on $G$ such that the group of isomorphism classes in this category is canonically identified with the $\ell$-adic character group of $G(\Fq)$, thus providing the geometrization and categorification of characters of $G(\Fq)$.
%
Although hidden in this discussion, the connectedness of $G$ played a crucial role in the geometrization and categorification of characters of $G(\Fq)$.
In particular, when the isomorphism \eqref{introbox} exists for $\qcs{L}$, it is essentially unique.

Now let $G$ be any smooth commutative group scheme over $\Fq$.
Two new features must be addressed in order to geometrize characters of $G(\Fq)$.
First, we must pass from $\ell$-adic local systems on $G$ to Weil local systems on $G$
and, second, we must keep track of the isomorphisms \eqref{introbox}.
Both issues are addressed in the main definition of this paper:
A \emph{quasicharacter sheaf} on $G$ is a triple $\qcs{L}\ceq
(\gqcs{L},\mu,\phi)$ where $\gqcs{L}$ is an $\ell$-adic local system on $\bG$ of rank~one and $\mu: m^*
\gqcs{L} \to \LxL$ and $\phi : \Frob{G}^*\gqcs{L} \to \gqcs{L}$ are isomorphisms satisfying natural compatibility conditions; here $\Frob{G}$ is the Frobenius automorphism of ${\bar G}$.
 In the main result of this paper, Theorem~\ref{thm:snake},
we form a rigid monoidal category $\lQC(G)$ of quasicharacter sheaves on $G$
and show that the Trace of Frobenius provides a functorial group isomorphism
\begin{equation}
\lQCiso{G} \to \Hom(G(\Fq), \EEx).
\end{equation}
\todo{change to remove flabbyness}

@@

%\subsection*{Structure of the paper}
\todo{Describe the overall structure of the paper in enough detail that one can follow the argument; include snake lemma diagram}

We close the introduction with a brief description of the structure of this paper.
After setting some basic notation in Section~\ref{sec:notation},
in Section~\ref{sec:category} we define quasicharacter sheaves and give an interpretation of
them as Weil sheaves equipped with an isomorphism of the form \eqref{introbox}.
Section~\ref{sec:Frob} introduces Trace of Frobenius,
one direction of the function--sheaf dictionary.
We define pullbacks and products of quasicharacter sheaves in Section~\ref{sec:pullback}, and use
them to prove a functoriality result on Trace of Frobenius.
In Section~\ref{sec:basechange} we study how quasicharacter sheaves behave under base change
and Weil restriction.
Sections~\ref{sec:bounded} and \ref{sec:finite} describe bounded quasicharacter sheaves and finite quasicharacter sheaves, respectively,
forming two important subcategories of $\QC(G)$.
%These subcategories play a key role in the proof of Proposition~\ref{prop:restriction}.
In Section \ref{sec:connected} we show that all quasicharacter sheaves on $G$ are finite quasicharacter sheaves when $G$ is connected, and that Trace of Frobenius is an isomorphism in this case.
We shift to \'etale group schemes $G$ in Section~\ref{sec:etale} and find a cohomological interpretation
of quasicharacter sheaves in this case.
In Section~\ref{sec:restriction} we show that for non-connected $G$, restriction to the identity component
induces an essentially surjective functor of quasicharacter sheaves.
%Section~\ref{sec:snake} presents the main theorem of this paper: Trace of Frobenius
%is a functorial isomorphism from isomorphism classes of flabby quasicharacter sheaves on $G$
%to characters of $G(\Fq)$.

\todo{Comments about Joseph Bernstein's comments?}

@@

Throughout this paper, $G$ is a smooth, commutative group scheme
over a finite field $\Fq$, and $m : \GxG\to G$ is its multiplication morphism.

@@

\subsection{The category of quasicharacter sheaves}\label{sec:category}

\begin{definition}\label{def:QC}
A \emph{quasicharacter sheaf on $G$} is a triple
$\qcs{L}\ceq (\gqcs{L},\mu,\phi)$ where:
\begin{enumerate}
\labitem{(QC.1)}{QC.1} $\gqcs{L}$ is a rank~one $\ell$-adic local system on $\bG$, by which we mean
a constructible $\ell$-adic sheaf on $\bG$, locally constant on each connected
component of $\bG$, whose stalks are one-dimensional $\EE$-vector spaces;
\labitem{(QC.2)}{QC.2} $\mu: \bm^* \gqcs{L} \to \LxL$ is an isomorphism of
sheaves on $\bGxG$ such that the following diagram commutes,
  where $m_3 \ceq m\circ (m\tight{1}{\times}{2}\id) = m\circ (\id\tight{2}{\times}{1} m)$;
  \[
  \begin{tikzcd}[row sep=30]
  \bm_3^*\gqcs{L} \arrow{rr}{(\bm \tight{1}{\times}{2} \id)^*\mu} \arrow[swap]{d}{(\id \tight{2}{\times}{1} \bm)^*\mu}
    &&  \bm^*\gqcs{L} \boxtimes \gqcs{L} \dar{\mu \tight{0}{\boxtimes}{1} \id} \\
    \gqcs{L} \boxtimes \bm^* \gqcs{L} \arrow{rr}{\id \boxtimes \mu}
    &&  \gqcs{L} \tight{0}{\boxtimes}{0} \LxL
  \end{tikzcd}
  \]
\labitem{(QC.3)}{QC.3} $\phi : \Frob{G}^* \gqcs{L} \to \gqcs{L}$ is an
  isomorphism of constructible $\ell$-adic sheaves on $\bG$ compatible with
  $\mu$ in the sense that the following diagram commutes.
  \[
  \begin{tikzcd}[row sep=20]
  \Frob{\GxxG}^* \bm^* \gqcs{L} \arrow{rr}{\Frob{\GxxG}^*\mu}
    && \Frob{\GxxG}^*(\LxL)\\
    \arrow[equal]{u} \bm^*  \Frob{G}^* \gqcs{L} \arrow[swap]{d}{\bm^* \phi}
    && \Frob{G}^*\gqcs{L}\boxtimes \Frob{G}^*\gqcs{L} \dar{\phi\boxtimes \phi} \arrow[equal]{u} \\
    \bm^*\gqcs{L} \arrow{rr}{\mu}
    && \LxL
  \end{tikzcd}
  \]
\end{enumerate}
\end{definition}

Morphisms in the category of quasicharacter sheaves on $G$, denoted by $\QC(G)$, are defined in the natural way:
\begin{enumerate}
\labitem{(QC.4)}{QC.4} if $\qcs{L} = (\gqcs{L},\mu,\phi)$ and
  $\qcs{L'} = (\gqcs{L'},\mu',\phi')$ are quasicharacter sheaves on $G$ then
  a morphism $\qcs{L} \to \qcs{L}'$ is a map $\alpha : \gqcs{L} \to \gqcs{L'}$
  of constructible $\ell$-adic sheaves on $\bG$ such that the following diagrams both commute.
  \[
  \begin{tikzcd}[column sep=40]
  \Frob{G}^* \gqcs{L} \rar{\Frob{G}^* \alpha} \arrow[swap]{d}{\phi} & \Frob{G}^* \gqcs{L'} \dar{\phi'}
  & & \arrow[swap]{d}{\mu} m^* \gqcs{L} \rar{m^* \alpha} & m^* \gqcs{L'} \dar{\mu'} \\
  \gqcs{L} \rar{\alpha} & \gqcs{L'}
  & {} & \LxL \rar{\tight{1}{\alpha\boxtimes \alpha}{1}} & \gqcs{L'} \tight{0}{\boxtimes}{0} \gqcs{L'}
  \end{tikzcd}
  \]
\end{enumerate}

The category $\QC(G)$ of quasicharacter sheaves on $G$ is a rigid monoidal category
\cite{etingof:09a}*{\S1.10} under the tensor product
$\qcs{L} \otimes \qcs{L'} \ceq (\gqcs{L}\otimes\gqcs{L'}, \mu\otimes\mu', \phi\otimes \phi')$,
with duals given by applying the sheaf hom functor
$\sheafHom(\ - \ ,\EE)$.
The category of quasicharacter sheaves on $G$ is not abelian, so $\QC(G)$
is not a tensor category in the sense of \cite{deligne:02a}*{0.1}.
This rigid monoidal category structure for $\QC(G)$ gives the set $\QCiso{G}$
of isomorphism classes in $\QC(G)$ the structure of a group.

%Observe that the rule $(\gqcs{L},\mu,\phi) \mapsto (\gqcs{L},\phi)$ defines a forgetful functor from quasicharacter
%sheaves on $G$ to ($\ell$-adic) Weil sheaves on $G$ \cite{deligne:80a}*{Def.~1.1.10 (i)}.
%

@@

@@

\begin{proposition}\label{prop:pullback}
  If $f : H\to G$ is a morphism of smooth commutative group schemes over $\Fq$ then
  \begin{align*}
  f^* : \QC(G) &\to \QC(H) \\
  (\gqcs{L},\mu,\phi) &\mapsto (\bf^*\gqcs{L},(\bfxf)^*\mu,\bf^*F)
  \end{align*}
  defines a monoidal functor dual to $f \colon H(\Fq) \to G(\Fq)$:
  \[
  \begin{tikzcd}[row sep=20, column sep=30]
   \QCiso{G} \rar{f^*} \arrow[swap]{d}{\TrFrob{G}} & \QCiso{H} \dar{\TrFrob{H}} \\
   G(\Fq)^* \rar & H(\Fq)^*
  \end{tikzcd}
  \]
  is a commutative diagram of groups.  Moreover, $(f\circ g)^* = g^* \circ f^*$.
\end{proposition}

@@

We end this section with a simple result on products of quasicharacter sheaves,
which is needed to show that the isomorphism above is functorial.

\begin{proposition}\label{prop:product}
If $G_1$ and $G_2$ are smooth commutative group schemes over $\Fq$ then the rule
$\boxtimes : (\qcs{L}_1,\qcs{L}_2)\to \qcs{L}_1\boxtimes\qcs{L}_2$ defines an equivalence of categories
\[
\QC(G_1)\times \QC(G_2) \to \QC(G_1\times G_2)
\]
such that
\[
\begin{tikzcd}[column sep=60]
\arrow{d}{\TrFrob{G_1} \times \TrFrob{G_2}} \QCiso{G_1}\times \QCiso{G_2} \arrow{r}{\boxtimes}
& \arrow{d}{\TrFrob{G_1\times G_2}} \QCiso{G_1\times G_2}\\
(G_1)(\Fq)^*\times (G_2)(\Fq)^* \arrow{r}{(\chi_1,\chi_2)\mapsto \chi_1\otimes \chi_2}  & (G_1\times G_2)(\Fq)^*
\end{tikzcd}
\]
commutes.
\end{proposition}
